Carryover effects in free recall reveal how prior experiences influence memories of new experiences

### 2023-01-12 paper/main.tex
added some additional stats for "random" conditions   - all of the additions involved comparing reduced (early) and reduced (late) to the reduced and feature rich conditions.

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the irrelevant visual features constant on early
lists, but allowed them to vary randomly on late lists. In a \textit{reduced
(late)} condition, we allowed the irrelevant visual features to vary randomly
on early lists, but held them constant on late lists. Given our above findings
that (a) participants tended to remember more words and exhibit stronger
clustering effects on feature rich (versus reduced) lists, and (b) participants
tended to remember more words and exhibit stronger clustering effects on early
(versus late) lists, we expected these differences to be enhance in the reduced
(early) condition and diminished in the reduced (late) condition. However, to
our surprise, participants in \textit{neither} condition exhibited reliable
early versus late differences in accuracy (reduced (early): $t(41) = 1.499, p =
0.141$; reduced (late): $t(40) = 1.462, p = 0.152$), temporal clustering
(reduced (early): $t(41) = 0.998, p = 0.324$; reduced (late): $t(40) = 1.099, p
= 0.278$), nor feature based clustering (reduced (early), category: $t(41) =
0.753, p = 0.456$; reduced (early), size: $t(41) = 0.721, p = 0.475$; reduced
(early), length: $t(41) = 0.493, p = 0.625$; reduced (early), first letter:
$t(41) = 0.780, p = 0.440$; reduced (late), category: $t(40) = -0.086, p =
0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$; reduced (late),
length: $t(40) = 1.476, p = 0.148$; reduced (late), first letter: $t(40) =
0.966, p = 0.340$). We hypothesized that adding or removing the irrelevant
features was acting as a sort of ``event boundary'' between early and late
lists. In prior work, we (and others) have found that memories formed just
after event boundaries can be enhanced~\citep[e.g.,][]{MannEtal16}.

@@

- are order effects "sticky"?  compare behavior on early vs. late lists for order manipulation condition
  - does feature clustering on early lists correlate with recall on early (or late) lists?
  - does feature clustering on late lists correlate with recall on early (or late) lists?
  - (ditto, but replace "recall" with "temporal clustering")

>>>>

With these differences in mind, we next compared participants' memories on
early versus late lists for two additional experimental conditions (see
\textit{Random conditions}, Fig.~\dynamicsRandom). In a \textit{reduced
(early)} condition, we held the irrelevant visual features constant on early
lists, but allowed them to vary randomly on late lists. In a \textit{reduced
(late)} condition, we allowed the irrelevant visual features to vary randomly
on early lists, but held them constant on late lists. Given our above findings
that (a) participants tended to remember more words and exhibit stronger
clustering effects on feature rich (versus reduced) lists, and (b) participants
tended to remember more words and exhibit stronger clustering effects on early
(versus late) lists, we expected these early versus late differences to be
enhanced in the reduced (early) condition and diminished in the reduced (late)
condition. However, to our surprise, participants in \textit{neither} condition
exhibited reliable early versus late differences in accuracy (reduced (early):
$t(41) = 1.499, p = 0.141$; reduced (late): $t(40) = 1.462, p = 0.152$),
temporal clustering (reduced (early): $t(41) = 0.998, p = 0.324$; reduced
(late): $t(40) = 1.099, p = 0.278$), nor feature based clustering (reduced
(early), category: $t(41) = 0.753, p = 0.456$; reduced (early), size: $t(41) =
0.721, p = 0.475$; reduced (early), length: $t(41) = 0.493, p = 0.625$; reduced
(early), first letter: $t(41) = 0.780, p = 0.440$; reduced (late), category:
$t(40) = -0.086, p = 0.932$; reduced (late), size: $t(40) = 0.746, p = 0.460$;
reduced (late), length: $t(40) = 1.476, p = 0.148$; reduced (late), first
letter: $t(40) = 0.966, p = 0.340$). We hypothesized that adding or removing
the irrelevant features was acting as a sort of ``event boundary'' between
early and late lists. In prior work, we (and others) have found that memories
formed just after event boundaries can be enhanced~\citep[e.g., due to less
contextual interference between pre- and post-boundary items;][]{MannEtal16}.

We found that \textit{adding} irrelevant visual features on later lists that
had not been present on early lists (as in the reduced (early) condition)
served to enhance recall performance relative to conditions where all lists had
the same blends of features (accuracy for feature rich versus reduced (early):
$t(107) = -2.230, p = 0.028$; reduced versus reduced (early): $t(101) = -2.045,
p = 0.043$; also see Fig.~\accuracyByList A). However, \textit{subtracting}
irrelevant visual features on later lists that \textit{had} been present on
early lists (as in the reduced (late) condition) did not appear to impact
recall performance (accuracy for feature rich versus reduced (late): $t(106) =
-0.638, p = 0.525$; reduced versus reduced (late): $t(100) = -0.407, p =
0.685$). These comparisons suggest that recall accuracy has a directional
component (i.e., accuracy is affected differently by removing features later
that had been present earlier versus adding features later that had
\textit{not} been present earlier). In contrast, we found that participants
exhibited more temporal and feature-based clustering when we added irrelevant
visual features to \textit{any} lists (comparisons of clustering on feature
rich and reduced lists are reported above; temporal clustering in reduced
versus reduced (early) and reduced versus reduced (late) conditions: $t$s $\leq
-9.780$, $p$s $< 0.001$; feature based clustering in reduced versus reduced
(early) and reduced versus reduced (late) conditions: $t$s $\leq -5.443$, $p$s
$< 0.001$). Temporal and feature-based clustering were not reliably different
in the feature rich, reduced (early), and reduced (late) conditions (temporal
clustering in feature rich versus reduced (early) and feature rich versus
reduced (late) conditions: $t$s $\geq -1.434$, $p$s $\geq 0.154$; feature based
clustering in feature rich versus reduced (early) and feature rich versus
reduced (late) conditions: $t$s $\geq -1.359$, $p$s $> 0.177$).

Taken together, our findings thus far suggest that adding item features that
change over time, even when they vary randomly and independently of the items,
can enhance participants' overall memory performance and can also enhance
temporal and feature-based clustering. To the extent that the number of item
features that vary from moment to moment approximates the ``richness'' of
participants' experiences, our findings suggest that participants remember
``richer'' stimuli better and organize richer stimuli more reliably in their
memories. Next, we turn to examine the memory effects of varying the temporal
ordering of different stimulus features while holding the features themselves
constant. We hypothesized that changing the order in which participants were
exposed to the words on a given list might enhance (or diminish) the relative
influence of different features. For example, presenting a set of words
alphabetically might enhance participants' attention to the studied items'
first letters, whereas sorting the same list of words by semantic category
might instead enhance participants' attention to the words' semantic
attributes. Importantly, we expected these order manipulations to hold even
when the variation in the total set of features (across words) was held
constant across lists (e.g., unlike in the reduced (early) and reduced (late)
conditions, where visual features were added or removed from a subset of the
lists participants studied).

@@

- when lists are sorted by a given feature, how is memory performance affected (relative to feature rich -- early lists only)?
  - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- early lists only

- are order effects "sticky"?
  - when *early* lists are sorted by a given feature, how is memory performance on *late* lists affected (relative to feature rich -- late lists only)?
  - do some order manipulations matter more than others?  e.g. compare semantic vs. lexicographic vs. visual -- late lists only

  - compare behavior on early vs. late lists for order manipulation condition
    - does feature clustering on early lists correlate with recall on early (or late) lists?
    - does feature clustering on late lists correlate with recall on early (or late) lists?
    - (ditto, but replace "recall" with "temporal clustering")
